### 2017-06-14 manuscript/manuscript.Rmd
Improve explanation of what VCS is

# The Git version control system

VCSs are computer programs designed for "tracking changes in computer files and coordinating work on those files among multiple people" (Wikipedia). In this tutorial, we focus on the most widely used (within science) VCS, called Git.

@@

It is important to understand that unlike an operating system's (OS) default file viewer, such as Finder (Apple computers) or File Explorer (?? Windows), Git is not a standalone program for navigating files and folders on a computer. Instead, it adds functionality to the OSs existing file system, by making available a specific set of functions--either executed from the command line, or through a graphical user interface (GUI) / integrated development environment (IDE)--that allow taking snapshots of the project and its files, and distributing the work across multiple computers and users. To begin using Git, users must first download and install the Git software on their computers. Git is free and open source, works on Windows, Macintosh, and Linux OSs (among others), and is used by major software developers such as Microsoft, Google, and Facebook. It can be freely downloaded at <https://git-scm.com/>.

>>>>

# Version Control Systems

Version Control Systems (VCS; also known as Source Control systems) are computer programs designed for tracking changes to computer code, and sharing the code---and its history---with others. Although initially developed for writing computer code collaboratively, it is easy to recognize the benefits of VCS for scientific production and collaboration as well. For example, behavioral experiments are often written in a programming language, and can include multiple authors and versions. Keeping track of the versions of the program, and allowing many authors to contribute to it (without breaking the code) are problems that VCSs were specifically designed for. By extension, VCS easily adopts to other aspects of the research cycle, such as curating data across computers, laboratories, and experiments; even writing manuscripts is easier when it is recognized that manuscripts (usually) undergo multiple rounds of revision, and that multiple authors might want to keep track of all these different versions.

The core concepts of version control are that contributors to a project create small checkpoints of the changes they make to the source code (or manuscript text, data, etc.), and then submit those changes to the VCS. The VCS maintains a history of all these little checkpoints, and the exact state of the project at each of these checkpoints.

@@

Some popular Version Control Systems are SVN, Mercurial, and Git. In this tutorial, we focus on the most widely used (in science) of these, called Git. Git is especially good for scientific collaboration because of online tools (GitHub) that allow seamless collaboration even for very large research teams.

## The Git Version Control System

It is important to understand that unlike an operating system's (OS) default file viewer, such as Finder (Apple computers) or File Explorer (?? Windows), Git is not a standalone program for navigating files and folders on a computer. Instead, it adds functionality to the OSs existing file system, by making available a specific set of functions--either executed from the command line, or through a graphical user interface (GUI) / integrated development environment (IDE)--that allow taking snapshots of the project and its files, and distributing the work across multiple computers and users. To begin using Git, users must first download and install the Git software on their computers. Git is free and open source, works on Windows, Macintosh, and Linux OSs (among others), and is used by major software developers such as Microsoft, Google, and Facebook. It can be freely downloaded at <https://git-scm.com/>. Before we explain the basics of using Git, we present brief installation instructions for Windows and Apple users.
